### 2025-04-03 paper/abs_en.tex
feat: paper

In the process of the transformation of modern agriculture towards intelligence, precise harvesting has become a crucial part in improving the efficiency and quality of agricultural production. Among them, the weighing of fruits and vegetables during the picking process is an important component of precise harvesting. With the widespread application of electronic scales, their role in the fruit and vegetable picking in smart farms has become increasingly prominent. However, at present, fruit and vegetable picking still largely relies on manual labor, which makes it particularly urgent to accurately record the weight of the picked fruits and scientifically manage the fruit and vegetable yields.

To meet the above - mentioned requirements, this study designed and implemented a cloud - based fruit weighing software with a front - end and back - end separation architecture based on the Java technology stack and the Vue framework. For the front - end, the Vue framework was selected for construction. Vue features a concise syntax, an efficient virtual DOM mechanism, and a good component - based development model, which can provide users with a smooth, aesthetically pleasing, and highly interactive operation interface. Users can conveniently configure electronic scales, view and manage data through this interface. The back - end adopts the Java technology stack, leveraging its rich class libraries and strong cross - platform capabilities to ensure the stability and scalability of the system. The Spring Boot framework is used to simplify the development process, and Spring JPA is employed for database operations, guaranteeing that the system can efficiently process a large amount of weighing data.

This software has multiple core functions. Firstly, it supports electronic scales based on different protocols such as HTTP and MQTT, enabling it to be compatible with various types of electronic scale devices on the market, which enhances the system's universality and adaptability. Secondly, it provides flexible configuration functions. Users can easily add new electronic scales, set corresponding keys, and configure data formats according to the data characteristics of different electronic scales, ensuring accurate parsing and processing of weighing data from various electronic scales. In addition, the software has powerful statistical analysis capabilities. It can conduct in - depth analysis of weighing data according to dimensions such as batches and pickers, export the analysis results in common data formats, and automatically generate detailed statistical reports, providing decision - making basis for farm managers.

To verify the functionality and performance of the software, this study simulated the interaction process between the weighing terminal and the cloud - based software. By simulating electronic scales with different protocols sending weighing data to the cloud - based software, a comprehensive test was carried out on the software's functionality integrity in data reception, parsing, storage, and processing. Meanwhile, the software performance was evaluated from multiple dimensions, including availability, Queries Per Second (QPS), concurrent user count, and response time. In terms of availability, long - term operation tests were conducted to ensure that the software can run stably and reliably under various conditions. The QPS test was used to evaluate the software's ability to process query requests per unit time. The concurrent user count test simulated scenarios where multiple users used the software simultaneously to examine the system's performance under high - concurrency situations. The response time test focused on the software's response speed after user operations to ensure a good user experience.

In conclusion, the cloud - based fruit weighing software implemented in this study provides an efficient, stable, and feature - rich solution for the weighing management of fruit and vegetable picking in smart farms, which contributes to promoting precise harvesting and digital management in smart farms.