### 2025-04-03 paper/main.tex

\documentclass{xduugthesis}

\usepackage{float} % \begin{figure}[H] 图片紧随文字
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{longtable}

\AtBeginDocument{\ctexset{subsection/format+=\centering}}
\xdusetup{
style = { 
    cjk-font = fandol, 
    latin-font = gyre,
    bib-backend = biblatex % -> biber
},
info = {
    title = {农业果实称重云端软件的设计与实现},
    department = {计算机科学与技术院},
    major = {计算机科学与技术},
    author = {江志航},
    supervisor = {霍秋艳},
    class-id = {2103018},
    student-id = {20009100359},
    abstract = {abs_zh.tex},
    abstract* = {abs_en.tex},
    keywords = {智慧农业, 称重系统, Spring Boot, Vue, MQTT, YOLO, 物联网, 信息管理系统},
    keywords* = {Smart agriculture, Weighing System, Spring Boot, Vue, MQTT, YOLO, IoT, Information management system},
    bib-resource = {reference.bib},
}
}
\begin{document}

\input{c1.tex}
\input{c2.tex}
\input{c3.tex}
\input{c4.tex}
\input{c5.tex}
\input{c6.tex}

\end{document}