### 2025-04-04 paper/abs_en.tex
feat: paper:c1

In the process of modern agriculture’s transformation towards intelligence, precision harvesting has become a key factor in improving agricultural production efficiency and quality. Among these, the weighing of fruits and vegetables during harvest is an essential part of precision harvesting. With the widespread application of electronic scales, their role in fruit and vegetable harvesting on smart farms has become increasingly prominent. However, at present, fruit and vegetable harvesting still largely depends on manual operations, making the accurate recording of harvested fruits’ weights and the scientific management of fruit yields particularly urgent.

Against this backdrop, this research conducts an in-depth survey of agricultural production activities, applying software engineering principles to identify and analyze specific requirements. It utilizes various software technologies, including web development techniques and YOLO-based fruit and vegetable recognition technologies, to design and develop a cloud-based agricultural fruit weighing software. This software achieves stable and high-performance data processing capabilities, provides efficient and rich information management functions, and offers multi-dimensional data visualization. Finally, the software’s various functionalities were tested, and performance tests on key interfaces were conducted in specific hardware environments, demonstrating the core indicators of the software to meet the deployment needs in practical production environments.

In conclusion, the cloud-based agricultural fruit weighing software developed in this study provides an efficient, stable, and feature-rich solution for smart farm fruit harvesting and weighing management, contributing to the advancement of precision harvesting and digital management in smart farms.