### 2025-05-07 paper/abs_en.tex
feat: polish15

As modern agriculture undergoes a transformation toward intelligent systems, the precise collection of fruit weighing data has become a critical foundation for efficient management and informed decision-making. However, traditional weighing often suffer from delayed data recording, frequent manual intervention, and poor hardware compatibility, which significantly hinder the efficiency and reliability of agricultural yield statistics.

To address these limitations, this study proposes and implements a cloud-based software system for agricultural fruit weighing that integrates multi-protocol communication, image recognition, and cloud data management. The system adopts a collaborative architecture between edge and remote nodes, ensuring independent data processing at both ends. At the edge, the system captures real-time fruit images and weight data, identifies fruit types using YOLO-based image recognition, and publishes weighing messages. The remote server subscribes to weighing data reported by the electronic scale and performs subsequent data processing and storage. The software supports multiple communication protocols such as MQTT, HTTP, and CoAP, enabling compatibility with various weighing devices for flexible integration and rapid deployment. A data management and visualization platform was developed using Spring Boot and Vue, offering efficient data management capabilities and multidimensional visualization of collected data. Comprehensive functional and performance tests were conducted under specific hardware environments, demonstrating the system’s ability to meet deployment requirements in real-world agricultural settings.

In conclusion, the proposed cloud-based fruit weighing software provides an efficient, robust, and feature-rich solution for intelligent fruit weighing and data management in smart farms, thereby facilitating precise harvesting and advancing the digital transformation of agriculture.